---
title: "Donut maps with DonutMap"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Donut maps with DonutMap}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

DonutMap creates donut charts positioned on a map from tidy data. The package
has three main workflows:

1. `donut_polygons()` creates an `sf` polygon layer.
2. `donut_map()` creates a static `ggplot2` map.
3. `donut_leaflet()` creates an interactive `leaflet` map with clickable
   donut segments and optional links or curved trajectories.

The examples below use a Natural Earth boundary loaded with `rnaturalearth`
(Natural Earth, n.d., <https://www.naturalearthdata.com/>). The point-level
category values and origin-destination trajectories are simulated for
demonstration.

```{r setup}
library(DonutMap)
library(ggplot2)
library(sf)

set.seed(20260522)
```

## Example data

The input data are tidy: each row gives a value for one category at one
location. Locations can be supplied either as longitude and latitude columns or
as an `sf` object.

```{r data}
sites <- data.frame(
  site = c("Site A", "Site B", "Site C", "Site D", "Site E"),
  lon = c(-73.57, -71.21, -72.75, -68.52, -66.82),
  lat = c(45.50, 46.81, 45.40, 48.45, 50.22)
)

categories <- c("Walking", "Transit", "Car")

demo <- merge(
  sites,
  data.frame(category = categories),
  by = NULL
)

demo$value <- c(
  32, 48, 120,
  55, 80, 95,
  28, 70, 110,
  20, 44, 76,
  18, 30, 58
)

demo$category <- factor(demo$category, levels = categories)

flows <- data.frame(
  from = c("Site A", "Site A", "Site B", "Site C", "Site D"),
  to = c("Site B", "Site C", "Site D", "Site E", "Site E"),
  trips = c(180, 90, 120, 70, 60)
)

category_colours <- c(
  Walking = "#1b9e77",
  Transit = "#7570b3",
  Car = "#d95f02"
)
```

For a background map, this vignette crops the Natural Earth Canada boundary to
eastern Canada. The donut values and flow values remain simulated.

```{r map-data}
canada <- rnaturalearth::ne_countries(
  country = "Canada",
  returnclass = "sf"
)

eastern_canada <- sf::st_crop(
  canada,
  sf::st_bbox(
    c(xmin = -81, ymin = 44, xmax = -62, ymax = 53.5),
    crs = sf::st_crs(4326)
  )
)
```

## Static map

`donut_map()` returns a normal `ggplot` object, so additional `ggplot2` layers,
scales, labels, and themes can be added afterwards. The `flows` argument adds
links between donut locations. Use `flow_curvature = 0` for straight links, and
larger positive or negative values for curved trajectories.

```{r static-map}
donut_map(
  demo,
  site,
  category,
  value,
  lon = lon,
  lat = lat,
  map = eastern_canada,
  crs = 3347,
  radius_range = c(25000, 70000),
  colours = category_colours,
  flows = flows,
  from = from,
  to = to,
  flow_value = trips,
  flow_linewidth_range = c(0.3, 2.2),
  flow_curvature = 0.22,
  flow_arrow = TRUE
) +
  labs(
    title = "Simulated mobility composition by site",
    fill = "Mode",
    linewidth = "Trips"
  ) +
  theme(legend.position = "right")
```

## Interactive map

`donut_leaflet()` returns a `leaflet` htmlwidget. Donut segments and trajectory
lines can be clicked to open popups, and hover labels are enabled by default.

```{r interactive-map}
donut_leaflet(
  demo,
  site,
  category,
  value,
  lon = lon,
  lat = lat,
  map = eastern_canada,
  crs = 3347,
  radius_range = c(25000, 70000),
  colours = category_colours,
  flows = flows,
  from = from,
  to = to,
  flow_value = trips,
  flow_weight_range = c(1, 7),
  flow_curvature = 0.22
)
```

## Trajectory geometries

`flow_lines()` returns the trajectory layer directly as an `sf` object. This is
useful when you want to build a custom map layer or inspect the geometry before
plotting.

```{r trajectory-layer}
trajectories <- flow_lines(
  flows,
  demo,
  from,
  to,
  trips,
  site,
  lon = lon,
  lat = lat,
  crs = 3347,
  flow_curvature = 0.22,
  flow_n = 40
)

trajectories
```

## Using the geometry layer directly

For more specialized workflows, `donut_polygons()` returns the donut segments as
valid `sf` polygons.

```{r sf-layer}
donuts <- donut_polygons(
  demo,
  site,
  category,
  value,
  lon = lon,
  lat = lat,
  crs = 3347,
  radius_range = c(25000, 70000)
)

donuts
```

You can use that object with any package that accepts `sf` polygons.

```{r sf-plot}
plot(donuts["category"])
```
